---
title: "Untitled"
author: "G1"
date: "6/6/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script takes the raw data from the CH2018 climate scenarios and outputs a raster in TIF format with the average projected highest daily temperature in the period between 2070 and 2099 for the four seasons.

```{r}
library(ncdf4) 
library(raster) 
library(mapview)
```

```{r}
# name for rcp 4.5
name <- "CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_"
# name for rcp 8.5
#name <- "CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP85_QMgrid_" 

# Load netcdf file
netcdf <- nc_open(paste0("raw_data/climate/future/raw_data/", name, "1981-2099.nc"))
```

```{r}
# Dates are given in "number of days since 01-01-1900" 
# --> we need to know which observations belong to winter, spring, summer and autumn in the period between 01-01-2070 - 31-12-2099
start_date <- as.Date("2070-01-01")
end_date <- as.Date("2099-12-31")
origin_date <- as.Date("1900-01-01")
origin_measurements <- as.Date(29585, origin="1900-01-01")

# Total measurement period
measurement_period <- seq(origin_measurements, end_date, by="day")

# Calculate the days in the period of interest
start_day <- as.integer(start_date - origin_date)
end_day <- as.integer(as.Date("2099-12-31") - origin_date)

dates_of_interest_num <- seq(start_day, end_day)
dates_of_interest <- as.Date(dates_of_interest_num, origin = "1900-01-01")

# Figuring out, which of these days belong to which season
winter_dates <- dates_of_interest[months(dates_of_interest) %in% c("December", "January", "February")]
spring_dates <- dates_of_interest[months(dates_of_interest) %in% c("March", "April", "May")]
summer_dates <- dates_of_interest[months(dates_of_interest) %in% c("June", "July", "August")]
autumn_dates <- dates_of_interest[months(dates_of_interest) %in% c("September", "October", "November")]

# What are the indices of these dates in the original measurement period?
winter_indices <- measurement_period %in% winter_dates
spring_indices <- measurement_period %in% spring_dates
summer_indices <- measurement_period %in% summer_dates
autumn_indices <- measurement_period %in% autumn_dates
```

After figuring out the time indices of the measurement values we need, we can filter the original NetCDF object and calculate the mean in the period to obtain one single output raster for every season.

```{r}
# extracting all temperature values, lat and lon values
temp_vals <- ncvar_get(netcdf, "tasmax")
lon_vals <- ncvar_get(netcdf, "lon")
lat_vals <- ncvar_get(netcdf, "lat")
```

```{r}
# extracting temperature values for every season
temp_winter <- temp_vals[,,winter_indices]
temp_spring <- temp_vals[,,spring_indices]
temp_summer <- temp_vals[,,summer_indices]
temp_autumn <- temp_vals[,,autumn_indices]

# Calculating mean per coordinate
mean_temp_winter <- as.matrix(rowMeans(temp_winter, dims=2))
mean_temp_spring <- as.matrix(rowMeans(temp_spring, dims=2))
mean_temp_summer <- as.matrix(rowMeans(temp_summer, dims=2))
mean_temp_autumn <- as.matrix(rowMeans(temp_autumn, dims=2))
```

```{r}
# CH1903 / LV95+
target_proj <- crs("+init=epsg:2056")

# creating a raster object and geocoding
mean_temp_winter_raster <- raster(t(mean_temp_winter[,ncol(mean_temp_winter):1]), xmn = min(lon_vals), xmx= max(lon_vals), ymn = min(lat_vals), ymx = max(lat_vals))
crs(mean_temp_winter_raster) <- "+proj=longlat +datum=WGS84"
mean_temp_winter_raster <- projectRaster(mean_temp_winter_raster, crs=target_proj)

mean_temp_spring_raster <- raster(t(mean_temp_spring[,ncol(mean_temp_spring):1]), xmn = min(lon_vals), xmx= max(lon_vals), ymn = min(lat_vals), ymx = max(lat_vals))
crs(mean_temp_spring_raster) <- "+proj=longlat +datum=WGS84"
mean_temp_spring_raster <- projectRaster(mean_temp_spring_raster, crs=target_proj)

mean_temp_summer_raster <- raster(t(mean_temp_summer[,ncol(mean_temp_summer):1]), xmn = min(lon_vals), xmx= max(lon_vals), ymn = min(lat_vals), ymx = max(lat_vals))
crs(mean_temp_summer_raster) <- "+proj=longlat +datum=WGS84"
mean_temp_summer_raster <- projectRaster(mean_temp_summer_raster, crs=target_proj)

mean_temp_autumn_raster <- raster(t(mean_temp_autumn[,ncol(mean_temp_autumn):1]), xmn = min(lon_vals), xmx= max(lon_vals), ymn = min(lat_vals), ymx = max(lat_vals))
crs(mean_temp_autumn_raster) <- "+proj=longlat +datum=WGS84"
mean_temp_autumn_raster <- projectRaster(mean_temp_autumn_raster, crs=target_proj)
```

```{r}
temperature_ch_july_rast <- raster("raw_data/climate/present/wald-lufttemperatur_juli_1981_2010_2056.tif")
temperature_ch_july_rast <- aggregate(temperature_ch_july_rast, fact=c(1000/25, 1000/25))

# resampling all the datasets to be the same resolution as the present climate data
mean_temp_winter_raster <- resample(mean_temp_winter_raster, temperature_ch_july_rast)
mean_temp_spring_raster <- resample(mean_temp_spring_raster, temperature_ch_july_rast)
mean_temp_summer_raster <- resample(mean_temp_summer_raster, temperature_ch_july_rast)
mean_temp_autumn_raster <- resample(mean_temp_autumn_raster, temperature_ch_july_rast)
```


```{r}
mapview(mean_temp_winter_raster)
```

```{r}
plot(mean_temp_summer_raster)
```


```{r}
# saving as TIFF file
writeRaster(mean_temp_winter_raster, filename= paste0("raw_data/climate/future/processed_data/", name, "2070-2099_winter.tif"), format="GTiff", overwrite=T)
writeRaster(mean_temp_spring_raster, filename= paste0("raw_data/climate/future/processed_data/", name, "2070-2099_spring.tif"), format="GTiff", overwrite=T)
writeRaster(mean_temp_summer_raster, filename= paste0("raw_data/climate/future/processed_data/", name, "2070-2099_summer.tif"), format="GTiff", overwrite=T)
writeRaster(mean_temp_autumn_raster, filename= paste0("raw_data/climate/future/processed_data/", name, "2070-2099_autumn.tif"), format="GTiff", overwrite=T)
```

```{r}
test <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_2070-2099_autumn.tif")

plot(test)
```

