---
title: "data_processing"
author: "G1"
date: "5/18/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(raster)
library(tidyverse)
library(sf)
library(leaflet)
library(data.table)
library(patchwork)
library(lubridate)
library(mapview)
library(rmapshaper)

options(dplyr.summarise.inform = FALSE)
```

### Creating the raster
```{r}
# Swiss plateau as basis
bio_geo_regions <- read_sf("raw_data/biogeographical_regions/N2020_Revision_BiogeoRegion.shp")

plateau <- bio_geo_regions %>%
  group_by(DERegionNa) %>%
  summarize(geometry=st_union(geometry)) %>%
  rename(name = DERegionNa) %>%
  filter(name == "Mittelland")

# Simplifying plateau geometry and saving as shp
#plateau_simp <- ms_simplify(plateau, keep_shapes=T, keep=0.005)
#st_write(plateau_simp, "processed_data/plateau/plateau.shp")
```

```{r warning=F}
# resolution of the hexagonal raster in m
resolution <- 2000

# Creating hexagonal fishnet based on geometry of plateau
target_fishnet <- st_sf(st_make_grid(plateau, cellsize=resolution, square=F)) %>%
  rename(geometry = st_make_grid.plateau..cellsize...resolution..square...F.)
target_fishnet <- target_fishnet[plateau,,]

# Adding row number for later join
target_fishnet <- mutate(target_fishnet, id = row_number())

# Generating points at centroid coordinates
target_fishnet_points <- target_fishnet %>%
  mutate(GKODE=st_coordinates(st_centroid(.))[,1]) %>%
  mutate(GKODN=st_coordinates(st_centroid(.))[,2]) %>%
  st_drop_geometry() %>%
  st_as_sf(., coords=c("GKODE", "GKODN"), crs=2056, agr="constant")

ggplot() +
  geom_sf(data=plateau) +
  geom_sf(data = target_fishnet, alpha=0.5) +
  theme_void()
```

```{r}
# creating a union of the fishnet for faster filtering purposes
target_fishnet_union <- target_fishnet %>%
  summarize(geometry = st_union(geometry))

#ggplot() +
#  geom_sf(data=plateau) +
#  geom_sf(data = target_fishnet_union, alpha=0.5) +
#  theme_void()
```

### Climate data

#### Present situation
```{r warning=F}
# reading the necessary raster data
temperature_ch_july_rast <- raster("raw_data/climate/present/wald-lufttemperatur_juli_1981_2010_2056.tif")
humidity_ch_july_rast <- raster("raw_data/climate/present/wald-relative_luftfeuchte_juli_1981_2010_2056.tif")

# downsample 25 m x 25 m --> 1 km x 1km
temperature_ch_july_rast <- aggregate(temperature_ch_july_rast, fact=c(1000/25, 1000/25))
humidity_ch_july_rast <- aggregate(humidity_ch_july_rast, fact=c(1000/25, 1000/25))

par(mfrow=c(1,2))
plot(temperature_ch_july_rast)
plot(humidity_ch_july_rast)
```

```{r}
## Temperature data

# Convert raster to points
points <- rasterToPoints(temperature_ch_july_rast)
points <- data.frame(gkode=points[,1], gkodn=points[,2], temp=points[,3])

# Create an sf object from the points
temperature_ch_july <- st_as_sf(points, coords = c("gkode", "gkodn"), crs = 2056)

# Assign raster values to sf object
temperature_ch_july$temp <- points[, "temp"]

# Remove all temperature points that are not in the Swiss plateau
temperature_ch_july <- temperature_ch_july[target_fishnet_union,,]

# assign hexagon cell to all temperature values in which they are contained
temperature_ch_july$id <- st_nearest_feature(temperature_ch_july, target_fishnet)

# compute mean temperature per hexagonal cell
setDT(temperature_ch_july)
temperature_ch_july_fishnet <- temperature_ch_july[, .(temp_july = mean(temp)), by = id]

# joining the data to the fishnet
target_fishnet <- left_join(target_fishnet, temperature_ch_july_fishnet, by="id")
```

```{r}
## Humidity data

# Convert raster to points
points <- rasterToPoints(humidity_ch_july_rast)
points <- data.frame(gkode=points[,1], gkodn=points[,2], hum=points[,3])

# Create an sf object from the points
humidity_ch_july <- st_as_sf(points, coords = c("gkode", "gkodn"), crs = 2056)

# Assign raster values to sf object
humidity_ch_july$hum <- points[, "hum"]

# Remove all humidity points that are not in the Swiss plateau
humidity_ch_july <- humidity_ch_july[target_fishnet_union,,]

# assign hexagon cell to all humidity values in which they are contained
humidity_ch_july$id <- st_nearest_feature(humidity_ch_july, target_fishnet)

# compute mean humidity per hexagonal cell
setDT(humidity_ch_july)
humidity_ch_july_fishnet <- humidity_ch_july[, .(hum_july = mean(hum)), by = id]

# joining the data to the fishnet
target_fishnet <- left_join(target_fishnet, humidity_ch_july_fishnet, by="id")
```

```{r}
temp <- ggplot() +
  geom_sf(data = plateau) +
  geom_sf(data = target_fishnet, aes(fill = temp_july), color=NA, alpha = 0.8) +
  scale_fill_gradient(low = "#FFCCCB", high = "#8b0000") +
  theme_void()

hum <- ggplot() +
  geom_sf(data=plateau) +
  geom_sf(data=target_fishnet, aes(fill = hum_july), color=NA, alpha=0.8)+
  scale_fill_gradient(low = "#FFCCCB", high="#90ee90") +
  theme_void()

temp + hum
```

We have the temperature and humidity levels for July, now we need to them for the remaining months, for which we use the data from the measurement stations located in the Swiss plateau over the same period.

```{r}
# Swiss meteo stations in plateau: Basel - Binningen, Bern - Zollikofen, Geneve / Cointrin, Luzern, Neuchatel, Payerne, Zurich / Fluntern, St. Gallen
stations_plateau <- c("BAS", "BER", "GVE", "LUZ", "NEU", "PAY", "SMA", "STG")

# locations of all Swiss meteo stations
stations <- read.csv("raw_data/climate/present/meteo_stations/meteoswissstations.csv")

stations <- stations %>%
  dplyr::filter(station.location %in% stations_plateau) %>%
  dplyr::select(`ï..Station`, station.location, CoordinatesE, CoordinatesN) %>%
  st_as_sf(., coords=c("CoordinatesE", "CoordinatesN"), crs=2056, agr="constant")

stations_mean_measurements <- data.frame()

# calculating average humidity and temperature for all stations for every months between 1981 and 2010
for (i in 1:length(stations_plateau)) {
  station_data <- read.csv(paste0("raw_data/climate/present/meteo_stations/nbcn-daily_",stations_plateau[i],"_previous.csv"), sep=";")
  
  # formatting date
  station_data$date <- ymd(station_data$date)
  station_data$year <- year(station_data$date)
  station_data$month <- month(station_data$date, label=T)
  
  # average daily max temperature and humidity per month between 1981 and 2010
  station_data_month_mean <- station_data %>%
    filter(year >= 1981 & year <= 2010) %>%
    group_by(month) %>%
    summarize(temp = mean(as.numeric(tre200dx)), hum=mean(as.numeric(ure200d0)))
  
  # pivoting the dataframe
  station_data_month_mean_pivot <- pivot_wider(
    data = station_data_month_mean,
    names_from = month,
    values_from = c(temp, hum)
  )
  
  # inserting station name into pivoted dataframe
  station_data_month_mean_pivot$station.location <- stations_plateau[i]
  
  stations_mean_measurements <- rbind(station_data_month_mean_pivot, stations_mean_measurements)
}

# joining the results to the stations sf object
stations <- left_join(stations, stations_mean_measurements, by="station.location")
```

Temperature and humidity values for months apart from July are calculated based on the delta between the respective of month and the July values for the closest station.

```{r}
# assigning closest station for every fishnet cell
target_fishnet$nearest_station <- stations_plateau[st_nearest_feature(target_fishnet, stations)]

# calculation of the missing month values
target_fishnet_measurements <- left_join(target_fishnet, stations %>% st_drop_geometry(), by=c("nearest_station" = "station.location")) %>%
  st_drop_geometry()

ggplot() +
  geom_sf(data = plateau) +
  geom_sf(data = target_fishnet, aes(fill = nearest_station), color=NA, alpha = 0.8) +
  theme_void()
```

```{r}
target_fishnet_measurements <- target_fishnet_measurements %>%
  mutate(temp_january = temp_july - (temp_Jul - temp_Jan)) %>%
  mutate(temp_february = temp_july - (temp_Jul - temp_Feb)) %>%
  mutate(temp_march = temp_july - (temp_Jul - temp_Mar)) %>%
  mutate(temp_april = temp_july - (temp_Jul - temp_Apr)) %>%
  mutate(temp_may = temp_july - (temp_Jul - temp_May)) %>%
  mutate(temp_june = temp_july - (temp_Jul - temp_Jun)) %>%
  mutate(temp_august = temp_july - (temp_Jul - temp_Aug)) %>%
  mutate(temp_september = temp_july - (temp_Jul - temp_Sep)) %>%
  mutate(temp_october = temp_july - (temp_Jul - temp_Oct)) %>%
  mutate(temp_november = temp_july - (temp_Jul - temp_Nov)) %>%
  mutate(temp_december = temp_july - (temp_Jul - temp_Dec)) %>%
  mutate(hum_january = hum_july - (hum_Jul - hum_Jan)) %>%
  mutate(hum_february = hum_july - (hum_Jul - hum_Feb)) %>%
  mutate(hum_march = hum_july - (hum_Jul - hum_Mar)) %>%
  mutate(hum_april = hum_july - (hum_Jul - hum_Apr)) %>%
  mutate(hum_may = hum_july - (hum_Jul - hum_May)) %>%
  mutate(hum_june = hum_july - (hum_Jul - hum_Jun)) %>%
  mutate(hum_august = hum_july - (hum_Jul - hum_Aug)) %>%
  mutate(hum_september = hum_july - (hum_Jul - hum_Sep)) %>%
  mutate(hum_october = hum_july - (hum_Jul - hum_Oct)) %>%
  mutate(hum_november = hum_july - (hum_Jul - hum_Nov)) %>%
  mutate(hum_december = hum_july - (hum_Jul - hum_Dec)) %>%
  
  # aggregating to seasons through means of the respective months
  mutate(temp_winter_present = rowMeans(dplyr::select(., temp_december, temp_january, temp_february))) %>%
  mutate(temp_spring_present = rowMeans(dplyr::select(., temp_march, temp_april, temp_may))) %>%
  mutate(temp_summer_present = rowMeans(dplyr::select(., temp_june, temp_july, temp_august))) %>%
  mutate(temp_autumn_present = rowMeans(dplyr::select(., temp_september, temp_october, temp_november))) %>%
  mutate(hum_winter_present = rowMeans(dplyr::select(., hum_december, hum_january, hum_february))) %>%
  mutate(hum_spring_present = rowMeans(dplyr::select(., hum_march, hum_april, hum_may))) %>%
  mutate(hum_summer_present = rowMeans(dplyr::select(., hum_june, hum_july, hum_august))) %>%
  mutate(hum_autumn_present = rowMeans(dplyr::select(., hum_september, hum_october, hum_november))) %>%
  
  # calculation of saturation deficit for every season according to Randolph & Storey (1999)
  mutate(sd_spring_present = (1 - (hum_spring_present/100)) * 4.9463 * exp(0.0621 * temp_spring_present)) %>%
  mutate(sd_summer_present = (1 - (hum_summer_present/100)) * 4.9463 * exp(0.0621 * temp_summer_present)) %>%
  mutate(sd_autumn_present = (1 - (hum_autumn_present/100)) * 4.9463 * exp(0.0621 * temp_autumn_present)) %>%
  mutate(sd_winter_present = (1 - (hum_winter_present/100)) * 4.9463 * exp(0.0621 * temp_winter_present))
```

```{r}
# rejoining to the geometry
target_fishnet <- target_fishnet %>%
  select(id, geometry) %>%
  left_join(target_fishnet_measurements %>% 
              dplyr::select(id, 
                            hum_winter_present, hum_spring_present, hum_summer_present, hum_autumn_present,
                            temp_winter_present, temp_spring_present, temp_summer_present, temp_autumn_present,
                            sd_winter_present, sd_spring_present, sd_summer_present, sd_autumn_present),
            by="id")
```

#### Future situation

```{r warning=F}
# Reading data processed in the script "climate_model_data_processing.rmd" for RCP 4.5 and 8.5 (all seasons)
temp_winter_45 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_2070-2099_winter.tif")
temp_spring_45 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_2070-2099_spring.tif")
temp_summer_45 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_2070-2099_summer.tif")
temp_autumn_45 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP45_QMgrid_2070-2099_autumn.tif")
temp_winter_85 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP85_QMgrid_2070-2099_winter.tif")
temp_spring_85 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP85_QMgrid_2070-2099_spring.tif")
temp_summer_85 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP85_QMgrid_2070-2099_summer.tif")
temp_autumn_85 <- raster("raw_data/climate/future/processed_data/CH2018_tasmax_CLMCOM-CCLM4_ECEARTH_EUR11_RCP85_QMgrid_2070-2099_autumn.tif")

# Function that computes the average temperature in each grid cell
computeAvgTemp <- function(scenario){
  # Convert raster to points
  points <- rasterToPoints(scenario)
  points <- data.frame(gkode=points[,1], gkodn=points[,2], temp=points[,3])
  
  # Create an sf object from the points
  temperature_ch <- st_as_sf(points, coords = c("gkode", "gkodn"), crs = 2056)
  
  # Assign raster values to sf object
  temperature_ch$temp <- points[, "temp"]
  
  # Remove all temperature points that are not in the Swiss plateau
  temperature_ch <- temperature_ch[target_fishnet_union,,]
  
  # assign hexagon cell to all temperature values in which they are contained
  temperature_ch$id <- st_nearest_feature(temperature_ch, target_fishnet)
  
  # compute mean temperature per hexagonal cell
  setDT(temperature_ch)
  temperature_ch_fishnet <- temperature_ch[, .(temp = mean(temp)), by = id]
  
  return(temperature_ch_fishnet)
}

target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_winter_45) %>% rename(temp_winter_45 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_spring_45) %>% rename(temp_spring_45 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_summer_45) %>% rename(temp_summer_45 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_autumn_45) %>% rename(temp_autumn_45 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_winter_85) %>% rename(temp_winter_85 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_spring_85) %>% rename(temp_spring_85 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_summer_85) %>% rename(temp_summer_85 = temp), 
                            by="id")
target_fishnet <- left_join(target_fishnet, 
                            computeAvgTemp(temp_autumn_85) %>% rename(temp_autumn_85 = temp), 
                            by="id")
```

```{r}
# Calculating saturation deficit for climate scenarios (assuming constant humidity)
target_fishnet <- target_fishnet %>%
  mutate(sd_winter_45 = (1 - (hum_winter_present/100)) * 4.9463 * exp(0.0621 * temp_winter_45)) %>%
  mutate(sd_spring_45 = (1 - (hum_spring_present/100)) * 4.9463 * exp(0.0621 * temp_spring_45)) %>%
  mutate(sd_summer_45 = (1 - (hum_summer_present/100)) * 4.9463 * exp(0.0621 * temp_summer_45)) %>%
  mutate(sd_autumn_45 = (1 - (hum_autumn_present/100)) * 4.9463 * exp(0.0621 * temp_autumn_45)) %>%
  mutate(sd_winter_85 = (1 - (hum_winter_present/100)) * 4.9463 * exp(0.0621 * temp_winter_85)) %>%
  mutate(sd_spring_85 = (1 - (hum_spring_present/100)) * 4.9463 * exp(0.0621 * temp_spring_85)) %>%
  mutate(sd_summer_85 = (1 - (hum_summer_present/100)) * 4.9463 * exp(0.0621 * temp_summer_85)) %>%
  mutate(sd_autumn_85 = (1 - (hum_autumn_present/100)) * 4.9463 * exp(0.0621 * temp_autumn_85))
```

### Land cover
```{r}
# Loading Arealstatistik
arealstatistik <- read.csv("raw_data/land_cover/ag-b-00.03-37-area-csv.csv", sep=";") %>%
  select(E, N, AS18_17) %>%
  rename(id_landcover = AS18_17) %>%
  st_as_sf(coords = c("E", "N"), crs=2056, agr = "constant")

arealstatistik_plateau <- arealstatistik[target_fishnet_union,,]

rm(arealstatistik)
```

```{r}
# Defining the weights for the 17 classes AS_17 defined in a separate csv
land_cover_classes <- read.csv("raw_data/land_cover/land_cover_weights.csv")

# Computing sum of weights per hexagon cell
arealstatistik_plateau <- left_join(arealstatistik_plateau, land_cover_classes, by="id_landcover")
arealstatistik_hexagon <- st_join(arealstatistik_plateau, target_fishnet, join = st_within)

arealstatstik_per_hexagon <- setDT(arealstatistik_hexagon)[, .(land_cover = sum(weight)), by = id]

target_fishnet <- left_join(target_fishnet, arealstatstik_per_hexagon, by="id") %>%
  mutate(land_cover = ifelse(is.na(land_cover), 0, land_cover))
```

```{r}
#mapview(target_fishnet, zcol="land_cover")
```

### Exposure
```{r}
# Calculating accessibility using the newer STATPOP dataset
bfs_pop <- read.csv("raw_data/statpop/STATPOP2021.csv", sep=";") %>%
  select(E_KOORD, N_KOORD, B21BTOT) %>%
  rename(pop = B21BTOT) %>%
  st_as_sf(coords = c("E_KOORD", "N_KOORD"), crs=2056, agr="constant")

# radius that is considered "accessible"
radius <- 15000

# removing data points outside plateau, buffer with radius to ensure calculation is correct at the edges
bfs_pop_plateau <- bfs_pop[target_fishnet_union %>% st_buffer(radius),,]

# calculating sum of population within given radius of every raster cell
hexagon_pop <- st_join(target_fishnet, bfs_pop_plateau, join = st_is_within_distance, dist = radius)

acc_pop_per_hexagon <- setDT(hexagon_pop)[, .(pop_acc = sum(pop)), by = id]
target_fishnet <- left_join(target_fishnet, acc_pop_per_hexagon, by="id")
```

```{r}
mapview(target_fishnet, zcol="pop_acc")
```

```{r}
# shortening column names
target_fishnet_abb <- target_fishnet %>%
  rename(h_w_2023 = hum_winter_present) %>%
  rename(h_sp_2023 = hum_spring_present) %>%
  rename(h_su_2023 = hum_summer_present) %>%
  rename(h_a_2023 = hum_autumn_present) %>%
  rename(t_w_2023 = temp_winter_present) %>%
  rename(t_sp_2023 = temp_spring_present) %>%
  rename(t_su_2023 = temp_summer_present) %>%
  rename(t_a_2023 = temp_autumn_present) %>%
  rename(sd_w_2023 = sd_winter_present) %>%
  rename(sd_sp_2023 = sd_spring_present) %>%
  rename(sd_su_2023 = sd_summer_present) %>%
  rename(sd_a_2023 = sd_autumn_present) %>%
  
  rename(t_w_45 = temp_winter_45) %>%
  rename(t_sp_45 = temp_spring_45) %>%
  rename(t_su_45 = temp_summer_45) %>%
  rename(t_a_45 = temp_autumn_45) %>%
  rename(sd_w_45 = sd_winter_45) %>%
  rename(sd_sp_45 = sd_spring_45) %>%
  rename(sd_su_45 = sd_summer_45) %>%
  rename(sd_a_45 = sd_autumn_45) %>%
  
  rename(t_w_85 = temp_winter_85) %>%
  rename(t_sp_85 = temp_spring_85) %>%
  rename(t_su_85 = temp_summer_85) %>%
  rename(t_a_85 = temp_autumn_85) %>%
  rename(sd_w_85 = sd_winter_85) %>%
  rename(sd_sp_85 = sd_spring_85) %>%
  rename(sd_su_85 = sd_summer_85) %>%
  rename(sd_a_85 = sd_autumn_85) %>%
  
  rename(lnd_cvr = land_cover)

# save fishnet with all intermediate variables as shp
#st_write(na.omit(target_fishnet_abb), "processed_data/fishnet/fishnet.shp", append=F)
```

### Aggregation
```{r}
fishnet <- read_sf("processed_data/fishnet/fishnet.shp")

# inserting column that is zero if temp < 7 and 1 if temp >= 7
fishnet <- fishnet %>%
  mutate(temp_above_7_winter_2023 = ifelse(t_w_2023 >= 7, 1, 0)) %>%
  mutate(temp_above_7_spring_2023 = ifelse(t_sp_2023 >= 7, 1, 0)) %>%
  mutate(temp_above_7_summer_2023 = ifelse(t_su_2023 >= 7, 1, 0)) %>%
  mutate(temp_above_7_autumn_2023 = ifelse(t_a_2023 >= 7, 1, 0)) %>%
  mutate(temp_above_7_winter_45 = ifelse(t_w_45 >= 7, 1, 0)) %>%
  mutate(temp_above_7_spring_45 = ifelse(t_sp_45 >= 7, 1, 0)) %>%
  mutate(temp_above_7_summer_45 = ifelse(t_su_45 >= 7, 1, 0)) %>%
  mutate(temp_above_7_autumn_45 = ifelse(t_a_45 >= 7, 1, 0)) %>%
  mutate(temp_above_7_winter_85 = ifelse(t_w_85 >= 7, 1, 0)) %>%
  mutate(temp_above_7_spring_85 = ifelse(t_sp_85 >= 7, 1, 0)) %>%
  mutate(temp_above_7_summer_85 = ifelse(t_su_85 >= 7, 1, 0)) %>%
  mutate(temp_above_7_autumn_85 = ifelse(t_a_85 >= 7, 1, 0))

# maximum saturation deficit value in a year
max_sd <- max(max(fishnet$sd_sp_2023), 
              max(fishnet$sd_su_2023), 
              max(fishnet$sd_a_2023),
              max(fishnet$sd_w_2023),
              max(fishnet$sd_sp_45),
              max(fishnet$sd_su_45),
              max(fishnet$sd_a_45),
              max(fishnet$sd_w_45),
              max(fishnet$sd_sp_85),
              max(fishnet$sd_su_85),
              max(fishnet$sd_a_85),
              max(fishnet$sd_w_85))

# inverting saturation deficit values by subtracting from the max (--> highest saturation deficit values become zero)
fishnet <- fishnet %>%
  mutate(sd_spring_2023_inv = max_sd - sd_sp_2023) %>%
  mutate(sd_summer_2023_inv = max_sd - sd_su_2023) %>%
  mutate(sd_autumn_2023_inv = max_sd - sd_a_2023) %>%
  mutate(sd_winter_2023_inv = max_sd - sd_w_2023) %>%
  mutate(sd_spring_45_inv = max_sd - sd_sp_45) %>%
  mutate(sd_summer_45_inv = max_sd - sd_su_45) %>%
  mutate(sd_autumn_45_inv = max_sd - sd_a_45) %>%
  mutate(sd_winter_45_inv = max_sd - sd_w_45) %>%
  mutate(sd_spring_85_inv = max_sd - sd_sp_85) %>%
  mutate(sd_summer_85_inv = max_sd - sd_su_85) %>%
  mutate(sd_autumn_85_inv = max_sd - sd_a_85) %>%
  mutate(sd_winter_85_inv = max_sd - sd_w_85)

# maximum inverted saturation deficit value in a year
max_inv_sd <- max(max(fishnet$sd_spring_2023_inv), 
                  max(fishnet$sd_summer_2023_inv), 
                  max(fishnet$sd_autumn_2023_inv),
                  max(fishnet$sd_winter_2023_inv),
                  max(fishnet$sd_spring_45_inv),
                  max(fishnet$sd_summer_45_inv),
                  max(fishnet$sd_autumn_45_inv),
                  max(fishnet$sd_winter_45_inv),
                  max(fishnet$sd_spring_85_inv),
                  max(fishnet$sd_summer_85_inv),
                  max(fishnet$sd_autumn_85_inv),
                  max(fishnet$sd_winter_85_inv))

# minimum inverted saturation deficit value in a year
min_inv_sd <- min(min(fishnet$sd_spring_2023_inv), 
                  min(fishnet$sd_summer_2023_inv), 
                  min(fishnet$sd_autumn_2023_inv),
                  min(fishnet$sd_winter_2023_inv),
                  min(fishnet$sd_spring_45_inv),
                  min(fishnet$sd_summer_45_inv),
                  min(fishnet$sd_autumn_45_inv),
                  min(fishnet$sd_winter_45_inv),
                  min(fishnet$sd_spring_85_inv),
                  min(fishnet$sd_summer_85_inv),
                  min(fishnet$sd_autumn_85_inv),
                  min(fishnet$sd_winter_85_inv))

# normalizing the values between zero and one
fishnet <- fishnet %>%
  mutate(sd_spring_2023_norm = (sd_spring_2023_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_summer_2023_norm = (sd_summer_2023_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_autumn_2023_norm = (sd_autumn_2023_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_winter_2023_norm = (sd_winter_2023_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  
  mutate(sd_spring_45_norm = (sd_spring_45_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_summer_45_norm = (sd_summer_45_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_autumn_45_norm = (sd_autumn_45_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_winter_45_norm = (sd_winter_45_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  
  mutate(sd_spring_85_norm = (sd_spring_85_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_summer_85_norm = (sd_summer_85_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_autumn_85_norm = (sd_autumn_85_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  mutate(sd_winter_85_norm = (sd_winter_85_inv - min_inv_sd) /(max_inv_sd - min_inv_sd)) %>%
  
  mutate(land_cover_norm = (lnd_cvr - min(lnd_cvr)) / (max(lnd_cvr) - min(lnd_cvr))) %>%
  mutate(pop_acc_norm = (pop_acc - min(pop_acc)) / (max(pop_acc) - min(pop_acc)))

# Multiplication of the normalized values to obtain final tick bite risk
tick_risk <- fishnet %>%
  mutate(tick_risk_spring_2023 = temp_above_7_spring_2023 * sd_spring_2023_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_summer_2023 = temp_above_7_summer_2023 * sd_summer_2023_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_autumn_2023 = temp_above_7_autumn_2023 * sd_autumn_2023_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_winter_2023 = temp_above_7_winter_2023 * sd_winter_2023_norm * land_cover_norm * pop_acc_norm) %>%
  
  mutate(tick_risk_spring_45 = temp_above_7_spring_45 * sd_spring_45_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_summer_45 = temp_above_7_summer_45 * sd_summer_45_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_autumn_45 = temp_above_7_autumn_45 * sd_autumn_45_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_winter_45 = temp_above_7_winter_45 * sd_winter_45_norm * land_cover_norm * pop_acc_norm) %>%
  
  mutate(tick_risk_spring_85 = temp_above_7_spring_85 * sd_spring_85_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_summer_85 = temp_above_7_summer_85 * sd_summer_85_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_autumn_85 = temp_above_7_autumn_85 * sd_autumn_85_norm * land_cover_norm * pop_acc_norm) %>%
  mutate(tick_risk_winter_85 = temp_above_7_winter_85 * sd_winter_85_norm * land_cover_norm * pop_acc_norm) %>%
  
  select(tick_risk_spring_2023, tick_risk_summer_2023, tick_risk_autumn_2023, tick_risk_winter_2023,
         tick_risk_spring_45, tick_risk_summer_45, tick_risk_autumn_45, tick_risk_winter_45,
         tick_risk_spring_85, tick_risk_summer_85, tick_risk_autumn_85, tick_risk_winter_85) %>%
  
  rename(risk_sp_23 = tick_risk_spring_2023) %>%
  rename(risk_su_23 = tick_risk_summer_2023) %>%
  rename(risk_a_23 = tick_risk_autumn_2023) %>%
  rename(risk_w_23 = tick_risk_winter_2023) %>%
  rename(risk_sp_45 = tick_risk_spring_45) %>%
  rename(risk_su_45 = tick_risk_summer_45) %>%
  rename(risk_a_45 = tick_risk_autumn_45) %>%
  rename(risk_w_45 = tick_risk_winter_45) %>%
  rename(risk_sp_85 = tick_risk_spring_85) %>%
  rename(risk_su_85 = tick_risk_summer_85) %>%
  rename(risk_a_85 = tick_risk_autumn_85) %>%
  rename(risk_w_85 = tick_risk_winter_85)
```

```{r}
mapview(tick_risk, zcol="risk_w_85")
```

```{r}
# save final tick bite risk as shp
st_write(tick_risk, "processed_data/tick_bite_risk/tick_bite_risk.shp", append=F)
```

```{r}
# Loading the final data
tick_bite_risk <- st_transform(read_sf("processed_data/tick_bite_risk/tick_bite_risk.shp"), crs=4326)

mapview(tick_bite_risk, zcol="risk_su_85")
```

# Reported tick bites

```{r}
tickbites <- read.csv("raw_data/tick_bites/20230611-tick-reports.csv", sep=";") %>% 
  st_as_sf(coords=c("Lon", "Lat"), crs=4326) %>% 
  st_transform(2056) %>%
  select(geometry, date, ID) %>%
  mutate(month=month(ymd_hms(date))) %>%
  mutate(year=year(ymd_hms(date))) %>%
  mutate(season = ifelse(month %in% c(12, 1, 2), "Winter",
                      ifelse(month %in% c(3, 4, 5), "Spring",
                             ifelse(month %in% c(6, 7, 8), "Summer", "Autumn"))))

# removing tick bites outside mittelland
tickbites_plateau <- tickbites[target_fishnet_union,,]

st_write(tickbites_plateau, "processed_data/tick_bites/tick_bites_plateau.shp", append=F)
```

